---
title: "Creating `skimr`"
author: "Team skimr"
date: "5/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation 

The motivation of this project was to create a frictionless approach to quickly viewing summary statistics as part of a pipeline. There are many existing summary functions, but we found them lacking in one way or another because they can be generic, they don't always provide easy-to-operate-on data structures, and they are not pipeable. 

So at [rOpenSci #unconf17](http://unconf17.ropensci.org/), we created a new package that would let you quickly skim useful, tidy summary statistics directly from a pipe. 

And so we created [`skimr`](https://github.com/ropenscilabs/skimr#skimr). 

In a nutshell, `skimr` will create a `skimr` object that can be further operated upon or that provides a human-readable printout in the console. It presents reasonable default summary statistics for numerics, factors, etc, and lists counts, and missing and unique values. 

## Dev process

### About Team Members

**Amelia McNamara**  
Job Title: Visiting Assistant Professor of Statistical & Data Sciences at Smith College  
Project Contributions: Coder

**Eduardo Arino de la Rubia**  
Job Title: Chief Data Scientist at Domino Data Lab    
Project Contributions: Coder

**Hao Zhu**  
Job Title: Programmer Analyst at the Institute for Aging Research  
Project Contributions: Coder

**Julia Lowndes**  
Job Title: Marine Data Scientist at the National Center for Ecological Analysis and Synthesis   
Project Contributions: Documention and test scripts

**Shannon Ellis**  
Job Title: Postdoctoral fellow in the Biostatistics Department at the Johns Hopkins Bloomberg School of Public Health  
Project Contributions: Test Scripts

**Elin Waring**  
Job Title: Professor at Lehman College Sociology Department, City University of New York  
Project Contributions: Coder

**Michael Quinn**  
Job Title: Quantitative Analyst at Google  
Project Contributions: Coder

**Hope McLeod**  
Job Title: Data Engineer at Kobalt Music  
Project Contributions: Documentation

- how we pulled it all together
- next steps/future (issues for later)

### Step-by-step thought process

We started off by brainstorming what we liked about existing summary packages and what other features we wanted. We started looking at example data, `mtcars`. 

```{r, eval=FALSE}
str(mtcars)
summary(mtcars)
```

#### What we like and dislike, in Amelia's words

```{r summary}
# "I like what we get here because mpg is numeric so these stats make sense:" 
summary(mtcars$mpg) 

# "But I don’t like this because cyl should really be a factor and shouldn't have these stats:"
summary(mtcars$cyl)

# "This is OK, but not descriptive enough. It could be clearer what I'm looking at."
mosaic::tally(~cyl, data=mtcars) # install.packages('mosaic')

# "But this output isn't labeled, not ideal." 
table(mtcars$cyl, mtcars$vs)

# "I like this because it returns 'sd', 'n' and 'missing'":
mosaic::favstats(~mpg, data=mtcars) 
```

## Introducing `skimr`



- what we've implemented to overcome those dislikes
- examples of skimr features

