---
title: "Creating `skimr`"
author: "Team skimr"
date: "5/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation 

in words, what we wanted from a summary function and why

## Dev process

- our team members/background in terms of coding
- what we each worked on/collaborated
- how we pulled it all together
- next steps/future (issues for later)

### Step-by-step thought process

We started off by brainstorming what we liked about existing summary packages and what other features we wanted. 

We started looking at example data, `mtcars`. 

```{r}
str(mtcars)
summary(mtcars)
```


### What we like and dislike, in Amelia's words

```{r summary}
# "I like what we get here because mpg is numeric so these stats make sense:" 
summary(mtcars$mpg) 

# "But I don’t like this because cyl should really be a factor and shouldn't have these stats:"
summary(mtcars$cyl)

# "This is OK, but not descriptive enough. It could be clearer what I'm looking at."
mosaic::tally(~cyl, data=mtcars) # install.packages('mosaic')

# "But this output isn't labeled, not ideal." 
table(mtcars$cyl, mtcars$vs)

# "I like this because it returns 'sd', 'n' and 'missing'":
mosaic::favstats(~mpg, data=mtcars) 
```

## Introducing `skimr`



- what we've implemented to overcome those dislikes
- examples of skimr features

