# skimr
```{r set-options, echo=FALSE}
devtools::dev_mode()
library(skimr)
skim_with_defaults()
options(tibble.width = Inf)
options(width = 120)
```

[![Build Status](https://travis-ci.org/ropenscilabs/skimr.svg?branch=master)](https://travis-ci.org/ropenscilabs/skimr)
[![codecov](https://codecov.io/gh/ropenscilabs/skimr/branch/master/graph/badge.svg)](https://codecov.io/gh/ropenscilabs/skimr)

The goal of skimr is to provide a frictionless approach to dealing with summary statistics iteratively and interactively as part of a pipeline, and that conforms to the principle of least surprise. 

`skimr` provides summary statistics that you can skim quickly to understand and your data and see what may be missing. It handles different data types (numerics, factors, etc), and returns a skimr object that can be piped or displayed nicely for the human reader. 

See our blog post [here](https://rawgit.com/ropenscilabs/skimr/master/blog.html).

## Installation


``` r
# install.packages("devtools")
devtools::install_github("hadley/pillar")
devtools::install_github("ropenscilabs/skimr")
```

## Skim statistics in the console

- added missing, complete, n, sd
- reports numeric/int/double separately from factor/chr
- handles dates, logicals
- uses [Hadley Wickham's pillar package](https://github.com/hadley/pillar), specifically `pillar::spark-bar()`

**Nicely separates numeric and factor variables:**  

```{r}
skim(chickwts)
```

**Many numeric variables:**  
```{r}
skim(mtcars)
```

 
**Another example:**  
```{r}
skim(iris)
```

## Handles grouped data

Skim() can handle data that has been grouped using `dplyr::group_by`.

```{r}
iris %>% dplyr::group_by(Species) %>% skim()
```

## skim_df object (long format)

By default `skim` prints beautifully in the console, but it also produces a long, tidy-format skim_df object that can be computed on. 

```{r}
a <-  skim(chickwts)
dim(a)

```


```{r}
print.data.frame(skim(chickwts))
```


## Compute on the full skim_df object
```{r}
skim(mtcars) %>% dplyr::filter(stat=="hist")
```

## Works with strings, lists and other column classes!

```{r}
skim(dplyr::starwars)
```


## Specify your own statistics

```{r}
funs <- list(iqr = IQR,
    quantile = purrr::partial(quantile, probs = .99))
  skim_with(numeric = funs, append = FALSE)
  skim_v(iris$Sepal.Length)
# Restore defaults
  skim_with_defaults()
```


```{r}

 funs <- list(iqr = IQR,
    quantile = purrr::partial(quantile, probs = .99))
  skim_with(numeric = funs, append = FALSE)
  skim_v(iris$Sepal.Length)
# Restore defaults
  skim_with_defaults()

```

## Limitations of current version

Currently the print methods are still in early stages of development. Printing is limited to numeric, character,
and factor data types. Therefore although additional types that are supported by skim() 
and skim_v() will not display with the default printing.  To view these you may view and manipulate the 
skim object.

At the moment in addition to the three types with print support complex, logical, Date, POSIXct, and ts classes
are supported with skim_v methods and the results are in the skim object.

We are also aware that both print.skim and print.data.frame (used for the skim object)  do not handle 
significant digits incorrectly.  

### Windows support for spark histograms

Windows cannot print the spark-histogram characters when printing a data-frame. For example, 
`"▂▅▇"` is printed as `"<U+2582><U+2585><U+2587>"`. This longstanding problem [originates in 
the low-level code](http://r.789695.n4.nabble.com/Unicode-display-problem-with-data-frames-under-Windows-td4707639.html) 
for printing dataframes. One workaround for showing these characters in Windows is to set the CTYPE part of your locale to Chinese/Japanese/Korean with `Sys.setlocale("LC_CTYPE", "Chinese")`. These values do show up by default when printing a data-frame created by `skim()` as a list (`as.list()`) or as a matrix (`as.matrix()`).

### Printing spark histograms and line graphs in knitted documents

Spark-bar and spark-line work in the console but may not work when you knit them to a specific document format.
The same session that produces a correctly rendered HTML document may produce an incorrectly rendered PDF, 
for example. This issue can generally be addressed by changing fonts to one with good building block (for
histogtams) and braille support (for line graphs).  For example, the open font "DejaVu Sans" from 
the `extra font` package supports these.  You may also want to try wrapping your results in `knitr::kable()`.

Displays in documents of different types will vary. For example, one user found that the font 
"Yu Gothic UI Semilight"  produced consistent results for Microsoft Word and Libre Office Write.

## Contributing

We welcome issue reports and pull requests including adding support for different variable classes.
