---
output: github_document
---
# skimr
```{r set-options, echo=FALSE}
library(skimr)
skim_with_defaults()
options(tibble.width = Inf)
options(width = 120)
```

[![Build Status](https://travis-ci.org/ropenscilabs/skimr.svg?branch=master)](https://travis-ci.org/ropenscilabs/skimr)
[![codecov](https://codecov.io/gh/ropenscilabs/skimr/branch/master/graph/badge.svg)](https://codecov.io/gh/ropenscilabs/skimr)
[![](https://badges.ropensci.org/175_status.svg)](https://github.com/ropensci/onboarding/issues/175)

`skimr` provides a frictionless approach to summary statistics which conforms
to the [principle of least surprise](https://en.wikipedia.org/wiki/Principle_of_least_astonishment),
displaying summary statistics the user can skim quickly to understand their
data. It handles different data types and returns a `skim_df` object which can
be included in a pipeline or displayed nicely for the human reader.

## Installation

The current released version of `skimr` can be installed from CRAN. If you wish
to install the current build of the next release you can do so using the
following:

``` r
# install.packages("devtools")
devtools::install_github("ropenscilabs/skimr")
```
The APIs for this branch should be considered reasonably stable but still
subject to change if an issue is discovered.

To install the version with the most recent changes that have not yet been
incorporated in the master branch (and may not be):

``` r
devtools::install_github("ropenscilabs/skimr", ref = "develop")
```
Do not rely on APIs from the develop branch.

## Skim statistics in the console

`skimr`:

- Provides a larger set of statistics than `summary()`, including missing,
complete, n, sd
- reports numeric/int/dbl types separately from factor/chr
- handles dates, logicals
- supports spark-bar and spark-line based on
[Hadley Wickham's pillar package](https://github.com/hadley/pillar).

**Nicely separates variables by class:**  

```{r}
skim(chickwts)
```

**Presentation is in a compact horizontal format:**  
```{r}
skim(iris)
```

**Individual columns of a data frame can be selected using tidyverse-style
selectors:**

```{r}
skim(iris, Sepal.Length, Petal.Length)
```

**Handles grouped data: **

`skim()` can handle data that has been grouped using `dplyr::group_by`.

```{r}
iris %>% dplyr::group_by(Species) %>% skim()
```

**Options for kable and pander: **

Enhanced print options are available by piping to `kable()` or `pander()`. 

## `skim_df` object (long format)

By default `skim()` prints beautifully in the console, but it also produces a
long, tidy-format `skim_df` object that can be computed on. 

```{r}
a <-  skim(chickwts)
dim(a)
```

```{r}
print.data.frame(skim(chickwts))
```

## Compute on the full `skim_df` object
```{r}
skim(mtcars) %>% dplyr::filter(stat=="hist")
```

## Works with strings, lists and other column classes

```{r}
skim(dplyr::starwars)
```

Users can add new classes.

## Specify your own statistics

```{r}
funs <- list(
  iqr = IQR,
  quantile = purrr::partial(quantile, probs = .99)
)

skim_with(numeric = funs, append = FALSE)
skim(iris, Sepal.Length)

# Restore defaults
skim_with_defaults()

```

## Limitations of current version

We are aware that there are issues with rendering the inline histograms and
line charts in various contexts, some of which are described below. 

### Support for spark histograms

There are known issues with printing the spark-histogram characters when
printing a data frame. For example, `"▂▅▇"` is printed as
`"<U+2582><U+2585><U+2587>"`. This longstanding problem [originates in
the low-level code](http://r.789695.n4.nabble.com/Unicode-display-problem-with-data-frames-under-Windows-td4707639.html) for printing dataframes; for example,
there are reports of this issue in Emacs ESS.

This means that while `skimr` can render the histograms to the console and in
`kable()`, it cannot in other circumstances. This includes:

* rendering a `skimr` data frame within `pander()`
* converting a `skimr` data frame to a vanilla R data frame, but tibbles render
  correctly

One workaround for showing these characters in Windows is to set the CTYPE part
of your locale to Chinese/Japanese/Korean with `Sys.setlocale("LC_CTYPE",
"Chinese")`. These values do show up by default when printing a data-frame
created by `skim()` as a list (`as.list()`) or as a matrix (`as.matrix()`).

### Printing spark histograms and line graphs in knitted documents

Spark-bar and spark-line work in the console, but may not work when you knit
them to a specific document format. The same session that produces a correctly
rendered HTML document may produce an incorrectly rendered PDF, for example.
This issue can generally be addressed by changing fonts to one with good
building block (for histograms) and Braille support (for line graphs). For
example, the open font "DejaVu Sans" from the `extrafont` package supports
these. You may also want to try wrapping your results in `knitr::kable()`.
Please see the vignette on using fonts for details.

Displays in documents of different types will vary. For example, one user found
that the font "Yu Gothic UI Semilight" produced consistent results for
Microsoft Word and Libre Office Write.

## Contributing

We welcome issue reports and pull requests, including potentially adding
support for commonly used variable classes. However, in general, we encourage
users to take advantage of skimr's flexibility to add their own customized
classes. Please see the [contributing](contributing.md) and
[conduct](conduct.md) documents.
