---
title: "Extending skimr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extending skimr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

The `skim()` function summarizes data types contained within data frames and
objects that have `as.data.frame()` methods to coerce them into data frames. It
comes with a set of default summary functions for a wide variety of data types,
but this is not comprehensive. 

Package authors (and advanced users) can add support for skimming their specific
non-data-frame objects in their packages, and they can provide different
defaults in their own summary functions. This will require including skimr as a
dependency.

# Skimming objects that are not coercible to data frames

This example will illustrate this by creating support for the `lm` object
produced by `lm()`. For any object this involves two required elements and one
optional element.  This is a simple example, but for other types of objects
there may be much more complexity

If you are adding skim support to a package you will also need to add `skimr` to
the list of imports. 

```{r}
library(skimr)
```

The `lm()` function produces a complex object with class "lm".

```{r}
results <- lm(weight ~ feed, data = chickwts)
class(results)
attributes(results)
```

There is no as.data.frame method for an `lm` object.

```{r, eval = FALSE}
as.data.frame(results)
#> Error in as.data.frame.default(results) :
#>  cannot coerce class ‘"lm"’ to a data.frame
```


Unlike the example of having a new type of data in a column of a simple data
frame (for which we would create a `sfl`) frame in the "Using skimr" vignette,
this is a different type of challenge: an object that we might wish to skim, but
that cannot be directly skimmed. Therefore we need to make it into an object
that is either a data frame or coercible to a data frame.

In the case of the lm object, the `model` attribute is already a data frame. So
a very simple way to solve the challenge is to skim `results$model` directly.

```{r}
skim(results$model)
```

This is works, but we could go one step further and create a new function for
doing this directly.

```{r}
skim_lm <- function(.data) {
  .data <- .data$model
  skimr::skim(.data)
}

lm(weight ~ feed, data = chickwts) %>% skim_lm()
```

If desired, a more complex function can be created.  For example, the lm object
also contains fitted values and residuals.  We could incorporate these in the
data frame.

```{r}
skim_lm <- function(.data, fit = FALSE) {
  .data <- .data$model
  if (fit) {
    .data <- .data %>%
      dplyr::bind_cols(
        fitted = data.frame(results$fitted.values),
        residuals = data.frame(results$residuals)
      )
  }
  skimr::skim(.data)
}
```


```{r}
skim_lm(results, fit = TRUE)
```

A second example of the need for a special function is with `dist` objects. The
`UScitiesD` dat set is an example of this.

```{r}
class(UScitiesD)
UScitiesD
```

A `dist` object is most often, as in this case, lower triange matrices of
distances, which can be measured in various ways.  There are many packages that
produce dist objects and/or take dist objects as inputs, including those for
cluster analysis and multidimensional scaling. 

A simple solution to this is to follow a similar design to that for `lm`
objects.

```{r}
skim_dist <- function(.data) {
  .data <- data.frame(as.matrix(.data))
  skimr::skim(.data)
}
```

However, this has the limitation of treating the dist data as though it is
simple numeric data.

What we might want to do instead, is to create a new class, for example,
"distance" that is specifically for distance data. This will allow it to have
its own `sfl` or skimr function list.

As handling gets more comples, rather than make a new function it can be more
powerful to define an `as.data.frame` S3 method for dist objects, which will
allow it to integrate with skimr more completely and uses to use the `skim()`
function directly. In a package you will want to export this.

```{r}
as.data.frame.dist <- function(.data) {
  .data <- data.frame(as.matrix(.data))

  .data[] <- lapply(.data, structure, class = "distance", nms = names(.data))
  .data
}

as.data.frame(UScitiesD)
```

However, until an `sfl` is created, `skimr` will not recognize the class and
fall back to treating the data as if it were character data.

```{r}
skim(UScitiesD)
```

The solution to this is to define an `sfl` (skimr function list) specifically
for the `distance` class.

## Defining sfl's for a package

`skimr` has an opinionated list of functions for each class (e.g. numeric,
factor)  of data. The core package supports many commonly used classes, but
there are many others. You can investigate these defaults by calling
`get_default_skimmer_names()`.

What if your data type, like `distance`, isn't covered by defaults? `skimr`
usually falls back to treating the type as a character, which isn't necessarily
helpful. In this case, you're best off adding your data type with `skim_with()`.

Before we begin, we'll be using the following custom summary statistics
throughout. These functions find the nearest and furthest other location for
each location. 

One thing that is important to be aware of when creating statistics functions
for skimr is that skimr largely uses tibbles rather than base data frames. This
means that many base operations do not work as epected.

```{r}
get_nearest <- function(column) {
  closest <- which.min(column[column != 0])
  cities <- attr(column, "nms")[column != 0]
  toString(cities[closest])
}

get_furthest <- function(column) {
  furthest <- which.max(column[column != 0])
  cities <- attr(column, "nms")[column != 0]
  toString(cities[furthest])
}
```

This function, like all summary functions used by `skimr` has two notable
features.

*  It accepts a vector as its single argument
*  it returns a scalar, or in R terminology, a vector of length 1.

There are a lot of functions that fulfill these criteria:

*  Existing functions from base, stats, or other packages,
*  lambda's created using the Tidyverse-style syntax
*  custom functions that have been defined in the `skimr` package
*  custom functions that you have defined.

Not fulfilling the two criteria can lead to some very confusing behavior within
`skimr`. Beware! An example of this issue is the base `quantile()` function in
default `skimr` percentiles are returned by using `quantile()` five 
times. In the case of these functions, there could be ties which would result in
returning vectors that have length greater than 1. This is handled by collapsing
all of the tied values into a single string. 

Notice, also, that in the case of distance data we may wish to exclude distances
of 0, which indicate the distance from a place to itself. In finding the minimum
our function looks only at the distance to other places.

There are at least two ways that you might want to customize skimr handling of a
special data type within a package or your own work.  The first is to create a
custom skimming function.

````{r}
skim_with_dist <- skim_with(
  distance = sfl(
    nearest = get_nearest,
    furthest = get_furthest
  )
)

skim_with_dist(UScitiesD)
````

The example above creates a new *function*, and you can call that function on
a specific column with `distance` data to get the appropriate summary 
statistics. The `skim_with` factory also uses the default skimrs for things 
like factors, characters, and numerics. Therefore our `skim_with_dist` is like
the regular `skim` function with the added ability to summarize `distance`
columns.

While this works for any data type and you can also include it within any 
package (assuming your users load skimr), there is a second, even better,
approach. To take full advantage of `skimr`, we'll dig a bit into its API.

## Adding new methods

`skimr` has a lookup mechanism, based on the function `get_skimmers()`, to
find default summary functions for each class. This is based on the S3 class
system. You can learn more about it in
[*Advanced R*](https://adv-r.hadley.nz/s3.html).

This requires that you add `skimr` to your list of dependencies.

To export a new set of defaults for a data type, create a method for the generic
function `get_skimmers`. Each of those methods returns an `sfl` (skimr
function list) This is the same list-like data structure used in the
`skim_with()` example above. But note! There is one key difference. When adding
a generic we also want to identify the `skim_type` in the `sfl`. You will
probably want to use `skimr::get_skimmers.distance()` but that will not work in
a vignette.

In a package you will want to export this.

````{r}
#' @importFrom skimr get_skimmers
#' @export
get_skimmers.distance <- function(column) {
  sfl(
    skim_type = "distance",
    nearest = get_nearest,
    furthest = get_furthest
  )
}
````

The same strategy follows for other data types.

*  Create a method
*  return an `sfl`
*  make sure that the `skim_type` is included.

Users of your package should load `skimr` to get the `skim()` function (although
you could import and reexport it). Once loaded, a call to
`get_default_skimmer_names()` will return defaults for your data types as well! 

```{r}
get_default_skimmer_names()
```

They will then be able to use `skim()` directly.

```{r}
skim(UScitiesD)
```


## Conclusion

This is a very simple example. For some packages the custom statistics
will likely  be much more complex. The flexibility of `skimr` allows you to
manage that.

Thanks to Jakub Nowosad, Tiernan Martin, Edzer Pebesma, Michael Sumner, and 
Kyle Butts for inspiring and helping with the development of this code.
