---
title: "Supporting additional objects"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Supporting additional objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Introduction

The `skim()` function summarizes data types contained within data frames. It
comes with a set of default summary functions for a wide variety of data types,
but this is not comprehensive. Package authors can add support for skimming
their specific data types in their packages, and they can provide different
defaults in their own summary functions.

This example will illustrate this by creating support for the `sf` object
produced by the  "sf: Simple Features for R" package. For any object this
involves two required elements and one optional element.

- experiment with interactive changes
- create methods to `get_skimmers` for different objects within this package
- if needed, define any custom statistics

If you are adding skim support to a package you will also need to add `skimr`
to the list of imports. Note that in this vignette the actual analysis will
not be run because that would require importing the `sf` package just for this
example.  However to run it on your own you can install `sf` and then run the
following code.  Note that code in this vignette was not evaluated when
rendering the vignette in order to avoid forcing installation of sf.

```{r, eval = FALSE}
library(sf)
nc <- st_read(system.file("shape/nc.shp", package = "sf"))
```

```
Linking to GEOS 3.7.2, GDAL 2.4.2, PROJ 5.2.0
Reading layer `nc' from data source 
`/path/to/library/sf/shape/nc.shp' using driver `ESRI Shapefile'
Simple feature collection with 100 features and 14 fields
geometry type:  MULTIPOLYGON
dimension:      XY
bbox:           xmin: -84.32385 ymin: 33.88199 xmax: -75.45698 ymax: 36.58965
epsg (SRID):    4267
proj4string:    +proj=longlat +datum=NAD27 +no_defs
```
```{r, eval=FALSE}
class(nc)
```
```
[1] "sf"         "data.frame"
```
Unlike the example of having a new type of data in a column of a simple data 
frame in the "Using skimr" vignette, this is a different type of object 
with special attributes .

## Experiment interactively

`skimr` has an opinionated list of functions for each class (e.g. numeric,
factor)  of data. The core package supports many commonly used classes,
but there are many others. You can investigate these defaults by calling
`get_default_skimmer_names()`.

What if your data type isn't covered by defaults? `skimr` usually falls
back to treating the type as a character, which isn't necessarily helpful. In
this case, you're best off adding your data type with `skim_with()`.

Before we begin, we'll be using the following custom summary statistic
throughout. It's a naive example, but covers the requirements of what we need.

```{r, eval = FALSE}
funny_sf <- function(x) {
  length(x) + 1
}
```

This function, like all summary functions used by `skimr` has two notable
features.

*  It accepts a vector as its single argument
*  It returns a scalar

There are a lot of functions that fulfill these criteria:

* existing functions from base, stats, or other packages,
* lambda's created using the Tidyverse-style syntax
* custom functions that have been defined in the `skimr` package
* custom functions that you have defined.

Not fulfilling the two criteria can lead to some very confusing behavior within
`skimr`. Beware! An example of this issue is the base `quantile()` function in
default `skimr` percentiles are returned by using `quantile()` five 
times.

Next, we create a custom skimming function. To do this, we need to think about
the many specific classes of data in the `sf` package.  The following example
will build  support for `sfc_MULTIPOLYGON`, but note that we'll have to
eventually think about `sfc_LINESTRING`, `sfc_POLYGON`, `sfc_MULTIPOINT` and
others if we want to fully support `sf`.

```{r, eval = FALSE}
skim_sf <- skim_with(
  sfc_MULTIPOLYGON = sfl(
    n_unique = n_unique,
    valid = ~ sum(sf::st_is_valid(.)),
    funny = funny_sf
  )
)
```

The example above creates a new *function*, and you can call that function on
a specific column with `sfc_MULTIPOLYGON` data to get the appropriate summary 
statistics.

```{r, eval = FALSE}
skim_sf(nc$geometry)
```
```
── Data Summary ────────────────────────
                           Values                      
Name                       structure(list(geometry =...
Number of rows             100                         
Number of columns          1                           
_______________________                                
Column type frequency:                                 
  sfc_MULTIPOLYGON         1                           
________________________                               
Group variables            None                        

── Variable type: sfc_MULTIPOLYGON ──────────────────────────────────────────────────────
  skim_variable n_missing complete_rate n_unique valid funny missing     n
1 geometry              0             1      100   100   101       0   100
```
Creating a function that is a method of the skim_by_type generic
for the data type allows skimming of an entire data frame that contains some 
columns of that type.

```{r, eval = FALSE}
skim_by_type.sfc_MULTIPOLYGON <- function(mangled, columns, data) {
  skimmed <- dplyr::summarize_at(data, columns, mangled$funs)
  build_results(skimmed, columns, NULL)
}
```

```{r, eval = FALSE}
skim(nc)
```

```
── Data Summary ────────────────────────
                           Values
Name                       nc    
Number of rows             100   
Number of columns          15    
_______________________          
Column type frequency:           
  factor                   2     
  numeric                  12    
  sfc_MULTIPOLYGON         1     
________________________         
Group variables            None  

── Variable type: factor ────────────────────────────────────────────────────────────────
  skim_variable n_missing complete_rate ordered n_unique top_counts                    
1 NAME                  0             1 FALSE        100 Ala: 1, Ale: 1, All: 1, Ans: 1
2 FIPS                  0             1 FALSE        100 370: 1, 370: 1, 370: 1, 370: 1

── Variable type: numeric───────────────────────────────────────────────────────
   skim_variable n_missing complete_rate      mean        sd        p0       p25
 1 AREA                  0             1     0.126    0.0492     0.042     0.091
 2 PERIMETER             0             1     1.67     0.482      0.999     1.32 
 3 CNTY_                 0             1  1986.     107.      1825      1902.   
 4 CNTY_ID               0             1  1986.     107.      1825      1902.   
 5 FIPSNO                0             1 37100       58.0    37001     37050.   
 6 CRESS_ID              0             1    50.5     29.0        1        25.8  
 7 BIR74                 0             1  3300.    3848.       248      1077    
 8 SID74                 0             1     6.67     7.78       0         2    
 9 NWBIR74               0             1  1051.    1433.         1       190    
10 BIR79                 0             1  4224.    5179.       319      1336.   
11 SID79                 0             1     8.36     9.43       0         2    
12 NWBIR79               0             1  1353.    1976.         3       250.   
         p50       p75      p100 hist 
 1     0.120     0.154     0.241 ▆▇▆▃▂
 2     1.61      1.86      3.64  ▇▇▂▁▁
 3  1982      2067.     2241     ▇▆▆▅▁
 4  1982      2067.     2241     ▇▆▆▅▁
 5 37100     37150.    37199     ▇▇▇▇▇
 6    50.5      75.2     100     ▇▇▇▇▇
 7  2180.     3936     21588     ▇▁▁▁▁
 8     4         8.25     44     ▇▂▁▁▁
 9   698.     1168.     8027     ▇▁▁▁▁
10  2636      4889     30757     ▇▁▁▁▁
11     5        10.2      57     ▇▂▁▁▁
12   874.     1407.    11631     ▇▁▁▁▁

── Variable type: sfc_MULTIPOLYGON──────────────────────────────────────────────
  skim_variable n_missing complete_rate n_unique valid funny
1 geometry              0             1      100   100   101
```

Sharing these functions within a separate package requires an export. 
The simplest way to do this is with Roxygen.

```{r, eval = FALSE}
#' Skimming functions for `sfc_MULTIPOLYGON` objects.
#' @export
skim_sf <- skim_with(
  sfc_MULTIPOLYGON = sfl(
    missing = n_missing,
    n = length,
    n_unique = n_unique,
    valid = ~ sum(sf::st_is_valid(.)),
    funny = funny_sf
  )
)

#' A skim_by_type function for `sfc_MULTIPOLYGON` objects.
#' @export
skim_by_type.sfc_MULTIPOLYGON <- function(mangled, columns, data) {
  skimmed <- dplyr::summarize_at(data, columns, mangled$funs)
  build_results(skimmed, columns, NULL)
}
```

While this works within any package, there is an even better approach in this
case. To take full advantage of `skimr`, we'll dig a bit into its API.

## Adding new methods

`skimr` has a lookup mechanism, based on the function `get_skimmers()`, to
find default summary functions for each class. This is based on the S3 class
system. You can learn more about it in
[*Advanced R*](https://adv-r.hadley.nz/s3.html).

To export a new set of defaults for a data type, create a method for the generic
function `get_skimmers`. Each of those methods returns an `sfl`, a `skimr`
function list. This is the same list-like data structure used in the
`skim_with()` example above. But note! There is one key difference. When adding
a generic we also want to identify the `skim_type` in the `sfl`.

```{r, eval = FALSE}
#' @importFrom skimr get_skimmers
#' @export
get_skimmers.sfc_MULTIPOLYGON <- function(column) {
  sfl(
    skim_type = "sfc_MULTIPOLYGON",
    n_unique = n_unique,
    valid = ~ sum(sf::st_is_valid(.)),
    funny = funny_sf
  )
}
```

The same strategy follows for other data types.

* Create a method
* return an `sfl`
* make sure that the `skim_type` is there

```{r, eval = FALSE}
#' @export
get_skimmers.sfc_POINT <- function(column) {
  sfl(
    skim_type = "sfc_POINT",
    n_unique = n_unique,
    valid = ~ sum(sf::st_is_valid(.))
  )
}
```

Users of your package should load `skimr` to get the `skim()` function. Once
loaded, a call to `get_default_skimmer_names()` will return defaults for your
data types as well!

```{r, eval = FALSE}
get_default_skimmer_names()
```

```
...
$sfc_MULTIPOLYGON
[1] "n_unique" "valid"    "funny" 

$sfc_POINT
[1] "n_unique" "valid"  
```
## Conclusion

This is a very simple example. For a package such as sf the custom statistics
will likely  be much more complex. The flexibility of `skimr` allows you to
manage that.

Thanks to Jakub Nowosad, Tiernan Martin, Edzer Pebesma and Michael Sumner for
inspiring and  helping with the development of this code.
