---
title: "Using Fonts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Fonts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(extrafont)

```


One of the features of skim is the production of spark graphs for numeric data.
However these graphs may not render properly because of lack of font support or for other
reasons. In your specific environment this may depend on the fonts installed and the operating
system and may occur for only specific types of documents such as PDF documents.
Skimr supports kable() and pander() formatted tables, which used in conjunction with 
fonts to render the spark graphs.

To produce spark histograms a font that supports building blocks must be used. To
support the spark line graphs a font with Braille support must be used. 

A further discussion of this (with examples) is available in the "Using fonts" template for
skimr.  If you are using RStudio you can open that template by opening a new markdown
file and selecting "From template" and then choosing it.  Alternatively this file 
is available inside the skimr folder or repository at 
inst/markdown/templates/fonts_in_skimr/skeleton/skeleton.Rmd. 
