---
title: "Using Skimr"
author: "Elin Waring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Skimr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Under revision for skimr v2.

`skimr` is designed to provide summary statistics about variables. It is
opinionated in its defaults, but easy to modify.

In base R, the most similar functions are `summary()` for vectors and data
frames and `fivenum()` for numeric  vectors:

```{r}
summary(iris)
```
```{r}
summary(iris$Sepal.Length)
```
```{r}
fivenum(iris$Sepal.Length)
```
```{r}
summary(iris$Species)
```

# The `skim` function

The core function of skimr is `skim()`. `skim()` is a S3 generic function, with
methods for data frames, grouped data frames and vectors. Like `summary()`,
`skim()`'s method for data frames presents results for every column; the
statistics it provides depend on the class of the variable.

## Skimming data frames

By design, the main focus of `skimr` is on data frames; it is intended to fit
well withiin a data [pipeline](http://r4ds.had.co.nz/pipes.html) and relies
extensively on [tidyverse](https://www.tidyverse.org/) vocabulary, which
focuses on data frames.

Results of `skim()` are *printed* horizontally, with one section per variable
type and one row per variable. Results are *returned* from `skim()` as a long
[tibble](http://tibble.tidyverse.org/) of class `skim_df`, with one row per
variable + summary statistic.

```{r}
library(skimr)
skim(iris)
```

This is in contrast to `summary.data.frame()`, which stores statistics in a
`table`. The distinction is important, because the `skim_df` object is pipeable
and easy to use for additional manipulation: for example, the user could select
all of the variable means, or all summary statistics for a specific variable.

```{r}
skim(iris) %>%
  dplyr::filter(variable == "Petal.Length")
skim(iris) %>%
  dplyr::select(type, variable, mean)

```

The `skim_df` object always contains 2  columns: 

- `variable`: name of the original variable
- `type`: class of the variable

It also contains additional columns for each unique statistic (across the variable)
types.

The `skim_df` contains one row for each variable in the data. Statistics
that are not applicable for a given type of data are set to NA.

`skim()` also supports grouped data. In this case, one additional column for
each grouping variable is added to the `skim_df` object.

```{r}
mtcars %>%
  dplyr::group_by(gear) %>%
  skim()
```

Individual columns from a data frame may be selected using tidyverse-style
selectors.

```{r}
skim(iris, Sepal.Length, Species)
```
```{r}
skim(iris, starts_with("Sepal"))

```

If an individual column is of an unsuppported class, it is treated as a
character variable with a warning.

## Skimming vectors

In skimr v2 skim will attempt to coerce non-data frames (such as vectors
and matrices) to data frames.

`skim()` also handles individual vectors that are not part of a data frame. For
example, the `lynx` data set is class `ts`.

```{r}
skim(lynx)
```


## Skimming matrices

`skimr` does not support skimming matrices directly but coerces
them to data frames meaning that the matrix columns become variables 
similarly to `summary.matrix()`). Three possible ways to handle matrices 
with `skim()` parallel the three  variations of the mean function for matrices.

```{r}
m <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), nrow = 4, ncol = 3)
m
colMeans(m)
rowMeans(m)
mean(m)

skim(m)    # Similar to summary.matrix and colMeans()
skim(t(m)) # Similar to rowMeans()
skim(c(m)) # Similar to mean()

```


```

## Alternatives to `skim()`

`skim()` for a data frame returns a wide data frame. This long data
frame is printed horizontally as a separate summary for each data type found in
the data frame, but the object itself is not transformed during the print. 

Three other functions are available that may prove useful as part of `skim()`
workflows:

- `skim_tee()` produces the same printed version as `skim()` but returns the
original, unmodified data frame. This allows for continued piping of the
original data.

```{r}
iris_setosa <- iris %>%
  skim_tee(iris) %>%
  dplyr::filter(Species == "setosa")

```

- `partition()` returns a named list of the wide data frames for each data
type. These data frames contain the formatted, character values, meaning that
they are most useful for display. In general, users will want to store the
results in an object for further handling. Alternatively, `yank()` selects
only the subtable for a specific type.

```{r}
iris %>% skim() %>% partition()
iris %>% skim() %>% yank("numeric")

```


- `to_long()` returns a single long data frame with columns variable, 
type, statistic and formatted. This is similar but not identical to
the skim_df object in skimr v1.

```{r}
iris %>% skim() %>% to_long()
```

# Modifying skim() with skim_with() and skim_with_defaults()

`skimr` is opinionated in its choice of defaults, but users can easily add to,
replace, or remove the statistics for a class.   

To add a statistic, create a named list for each class using the format below:

```
#classname = list(mad_name = mad)

```

```{r}
#skim_with(numeric = list(mad_name = mad))
#skim(datasets::chickwts)

```

_When `skim_with()` is used to modify the statistics, the new list(s) of
statistics remains in place until they are reset using `skim_with_defaults()`._

By default `skim_with()` appends the new statistics, but setting
`append = FALSE` *replaces* the defaults.

```{r}
#skim_with_defaults()
#skim_with(numeric = list(mad_name = mad), append = FALSE)
#skim(datasets::chickwts)
#skim_with_defaults() # Reset to defaults

```

You can also use `skim_with()` to remove specific statistics by setting them to
`NULL`.

```{}
skim_with(numeric = list(hist = NULL))
skim(datasets::chickwts)
skim_with_defaults()

```

# Formatting individual values

When printing, `skimr` formats displayed statistics in an opinionated way;
these values are stored in the `formatted` column of the `skim_df` object and
are always character. `skim()` attempts to use a reasonable number of decimal
places for calculated values based on the data type (integer or numeric) and
number of stored decimals. For statistics such as `p0` and `p100`, the actual
stored values are displayed. Decimals in a column are aligned. Date formats are
used for date statistics. 

Users can override these opinionated formats using `skim_format()`.
`show_formats()` will display the current options in use for each data type.
Using `skim_format_defaults()` will reset the formats to their default
settings.

# Rendering the results of `skim()`

The `skim_df` object is a long data frame with one row for each combination of
variable and statistic (and optionally for group). The horizontal display is
created by default using `print.skim_df()`; users can specify additional
options by explicitly calling `print([skim_df object], ...)`.




## Solutions to common rendering problems

The details of rendering are dependent on the operating system R is running on,
the locale of the installation, and the fonts installed. Rendering may also
differ based on whether it occurs in the console or when knitting to specific
types of documents such as HTML and PDF.

The most commonly reported problems involve rendering the spark graphs (inline histogram) on Windows.

In order to render the sparkgraphs in html or PDF histogram you may need to
change fonts to one that supports blocks or Braille (depending on which you
need). Please review the separate vignette and associated template for details.
