---
title: "Using Skimr"
author: "Elin Waring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Skimr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The skimr package is designed to provide summary statistics about variables. In
base R the most similar functions are `summary()` for vectors and data frames
and `fivenum()` for numeric  vectors. Skimr is opinionated in its defaults but
easy to modify.

For comparison purposes here are examples of the similar functions.

```{r}
summary(iris)
```
```{r}
summary(iris$Sepal.Length)
```
```{r}
fivenum(iris$Sepal.Length)
```
```{r}
summary(iris$Species)
```

# The skim function

The core function of skimr is skim().  Skim is a S3 generic function; the skim
package  includes support for data frames and grouped data frames. Like summary
for data frames, skim presents results for all of the columns and the statistics
depend on the class of the variable.

However, unlike summary.data.frame(), the printed results (those displayed in
the console or in a knitted markdown file) are shown horizontally with one row
per variable and separated into separate tibbles for each class of variables.
The actual results are  stored in a `skim_df` object that is also a tibble 
In `summary.data.frame()` the statistics are stored in a `table` with
one column for each variable and the standard table printing is used to display
the results. 

```{r}
library(skimr)
skim(iris)
```

This distinction is important because the skim_df object is easy to use for
additional manipulation if desired and is pipeable. For example all of the
results for a particular statistic or for one  variable could be selected or
an alternative printing method sould be developed.  

The `skim_df` object always contains 6 columns: 

- `variable` the name of the original variable
- `type` the class of the variable
- `stat` the statistic calculated with the name becoming the column name when
  the object is printed.
- `level` is used when summary functions returns multiple values when skimming.
  This happens as for counts of levels for factor variables or when setting
  multiple values to the `probs` argument of the `quantiles` function.
- `value` is the actual calculated value of the statistic and should be used
  for further calculations. This is always numeric.
- `formatted` is a formatted version of value that attempts to use
  reasonable number of digits (decimal aligned) and put items such as dates into 
  human readable formats. It is a character variable.
 

```{r}
s <- skim(iris)
head(s, 15)
```


`skim()` also supports grouped data.  For grouped data one additional column for
each grouping  variable is added to the skim object.

```{r}
mtcars %>%
  dplyr::group_by(gear) %>%
  skim()
```

Individual columns from a data frame may be selected using tidyverse style selectors.

```{r}

skim(iris, Sepal.Length, Species)
skim(iris, starts_with("Sepal"))

```

If an individual column is of an unsuppported class it is treated as a 
character variable with a warning.

# Alternatives to skim

The skim() function for a data frame returns a long, six column data frame.
This long data frame is printed horizontally as a separate summary for each 
data type found in the data frame, but the object itself is not transformed 
during the print. 

Three other functions are available that may prove useful as part of skim 
work flows.

The `skim_tee()` function produces the same printed version as `skim()` but 
returns the unmodified data frame. This allows for continued piping of the 
original data.

The `skim_to_list()` funtion returns of a list of the wide data frames for 
each data type. The data frames contain the formatted values, meaning that 
they are character data and  most useful for display. In general users will 
want to store the results in an object for further handling.

The `skim_to_wide()` function returns a single data frame with each variable 
in a row. Variables that do not report a given statistic are assigned NA for 
that statistic. Formatted values are returned and all data are character.

# Skimming vectors

The `skim` function also handles individual vectors that are not part of a 
data frame. For example the `lynx` data set is class `ts`.

```{r}
skim(datasets::lynx)
```

If you attempt to use `skim` on a class that does not have support, it will
coerce it to  character (with a warning) and report number of `NA`s, number 
complete (non missing), number of rows, the  number empty (i.e. ""), minimum 
length of non empty strings, maximum length of non empty strings, and number 
of unique values.

```{r}
lynx <- datasets::lynx
class(lynx) <- "unkown_class"
skim(lynx)
```

# The skim_with function

Skimr is opinionated in its choice of defaults, but users can easily add too,
replace, or remove the statistics for a class.   

To add a statistic use the a named list for each class using the format below. 

```
classname = list(mad_name = mad)
```

```{r}
skim_with(numeric = list(mad_name = mad))
skim(datasets::chickwts)
```

The `skim_with_defaults()` function resets the list to the defaults. By default
`skim_with()` appends the new statstics, but setting `append = FALSE` replaces
the defaults. 

```{r}
skim_with_defaults()
skim_with(numeric = list(mad_name = mad), append = FALSE)
skim(datasets::chickwts)
skim_with_defaults() # Reset to defaults
```

You can also use `skim_with()` to remove specific statistics by setting them to
`NULL`.

```{r}
skim_with(numeric = list(hist = NULL))
skim(datasets::chickwts)
skim_with_defaults() #
```

# Formatting individual values

Skimr does opinionated formatting of the statistics displayed when printing.
These values are stored in the `formatted` column of the skim_df object and
are always character. Skim attempts to use a reasonable number of decimal 
places for calculated values based on the data type (integer or numeric) 
and number of stored decimals. For statistics such as `max()` and `min()` 
the actual stored values are displayed. Decimals in a column
are aligned.  Date formats are used for date statistics. 

Users override the formats using the `skim_format()` function. Using 
`show_formats()` will display the current options in use for each data
type. Using `skim_format_defaults()` will reset the formats to their default
settings.

# Rendering the results of skim()

The skim_df object is a long data frame with one row for each combination of 
variable and statistic (and optionally for group). 
The horizontal display is created by default using `print.skim_df()`. 
This can be called explicitly by applying the `print()`
function to a `skim_df` object which allows passing in of options. In addition
`kable()` and`pander()`are supported.  These both provide more control over
the rendered results, particularly when used to render in conjunction with knitr.
Documentation of these options for these functions is covered in more detail
in the `knitr` package for `kable()` and the `pander` package for `pander()`.
Using either of these may require use of document or chunk options and fonts.

* For `kable()` use in a markdown file, use a chunk option of `results='asis'`.
* For `pander()` use in a markdown file, use a chunk option of `results='asis'`. 
To prevent using asis by default use `panderOptons()` to set it to FALSE.

This topic is addressed in more detail in the Using Fonts vignette.

```{r results='asis'}

skim(iris) %>% kable()

```


```{r }
library(pander)
panderOptions('knitr.auto.asis', FALSE)
skim(iris) %>% pander() 

```


## Solutions to common rendering problems

The details of rendering are dependent on the operating system R is running on, 
the locale of the installation, and the fonts installed. Rendering may also
differ based on whether it occurs in the console or when knitting to specific
types of documents such as HTML and PDF. The most commonly reported problems
involve rendering the spark graphs (inline histogram and line graphs). 
This section will summarize known issues.  

Windows cannot print the spark-histogram characters when printing a data-frame. For example, 
`"▂▅▇"` is printed as `"<U+2582><U+2585><U+2587>"`. This longstanding problem [originates in 
the low-level code]
(http://r.789695.n4.nabble.com/Unicode-display-problem-with-data-frames-under-Windows-td4707639.html) 
for printing dataframes. One workaround for showing these characters in Windows is to set the
CTYPE part of your locale to Chinese/Japanese/Korean with 
`Sys.setlocale("LC_CTYPE", "Chinese")`. These values do show up by default when printing a
data-frame created by `skim()` as a list (`as.list()`) or as a matrix (`as.matrix()`).

In order to render the spark graphs in html or PDF histogram you may need to change
fonts to one that supports blocks or braille (depending on which you need). 
Please review the separate vignette and associated template for details on this.


