---
title: "Using Skimr"
author: "Elin Waring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Skimr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The skimr package is designed to provide summary statistics about variables in base R the
most similar functions are summary() for vectors and data frames and fivenum() for numeric 
vectors. It is opinionated in its defaults but easy to modify.

For comparison purposes here are examples of the similar functions.

```{r}
summary(iris)
```
```{r}
summary(iris$Sepal.Length)
```
```{r}
fivenum(iris$Sepal.Length)
```
```{r}
summary(iris$Species)

```


# The skim function

The core function of skimr is skim().  Skim is a generic function, but the skim package only 
includes support for data frames and grouped data frames. Like summary for data frames, skim
presents results for all of the columns and the statistics depend on the class of the variable.

However, unlike summary.data.frame(), the printed results (those displayed in the console
or in a knitted markdown file) are shown horizontally with one row per variable
and separated into separate tibbles for each class of variables.  The actual results are 
stored in a `skim_df` object that is also a tibble 
In `summary.data.frame()` the statistics are stored in a `table` with
one column for each variable and the standard table printing is used to display the results. 

```{r}
library(skimr)
skim(iris)
```

This distinction is important because the skim_df object is easy to use for additional manipulation
if desired and is pipeable. For example all of the results for a particular statistic or for one 
variable could be selected or an alternative printing method sould be developed.  

The skim_df object always contains 6 columns: 

 - `var` the name of the original variable
 - `type` the class of the variable
 - `stat` the statistic calculated with the name becoming the column name when the object is printed.
 - `level` is used when a statistic is referenced multiple times in the skim_df, such as for counts of levels for factor variables or the specific quantiles included.
 - `value` is the actual calculated value of the statistic and should be used for further calculations. This is always numeric.
 - `formatted_value` is a formatted version of value that attempts to use reasonable number of digits and put items such as dates into human readable formats. It is a character variable.
 

```{r}
s <- skim(iris)
head(as.data.frame(s), 15)
```


# The skim_v() function

The `skim_v()` function is the actual workhorse of `skimr` in that it processes each of the columns 
as a vector.  If you have data that is not a data frame one simple solution is to use `skim_v()` 
on individual vectors. For example the lynx data set is class `ts`.

```{r}
skim_v(datasets::lynx)
```


If you attempt to use `skim_v()` on a class that does not have support, it will coerce it to 
character (with a warning) and report number NA, number complete (non missing), number of rows, the 
number empty (i.e. ""), minimum length of non empty strings, maximum length of non empty strings,
and number of unique values.
```{r}
lynx <- datasets::lynx
class(lynx) <- "unkown_class"
skim_v(lynx)
```

# The skim_with() function

Skimr is opinionated in its choice of defaults, but users can easily add too, replace, or remove
the statistics for a class.   

To add a statistic use the a named list for each class using the format 
```
classname = list(mad_name = mad)
```
```{r}
skim_with(numeric = list(mad_name = mad))
skim(datasets::chickwts)
```

The `skim_with_defaults()` function resets the list to the defaults.   By default `skim_with()`
appends the new statstics, but setting append=FALSE replaces the defaults. 


```{r}
skim_with_defaults()
skim_with(numeric = list(mad_name = mad), append = FALSE)
skim(datasets::chickwts)
skim_with_defaults() # Reset to defaults
```
You can also use `skim_witn()` to remove specific statistics by setting them to NULL.

```{r}
skim_with(numeric = list(hist = NULL))
skim(datasets::chickwts)
skim_with_defaults() #
```


