---
output: 
  html_document: 
    keep_md: yes
---
# skimr
```{r set-options, echo=FALSE}
devtools::dev_mode()
library(skimr)
skim_with_defaults()
options(tibble.width = Inf)
options(width = 120)
```

[![Build Status](https://travis-ci.org/ropenscilabs/skimr.svg?branch=master)](https://travis-ci.org/ropenscilabs/skimr)
[![codecov](https://codecov.io/gh/ropenscilabs/skimr/branch/master/graph/badge.svg)](https://codecov.io/gh/ropenscilabs/skimr)

The goal of skimr is to provide a frictionless approach to dealing with summary statistics iteratively and interactively as part of a pipeline, and that conforms to the principle of least surprise. 

`Skimr` provides summary statistics that you can skim quickly to understand and your data and see what may be missing. It handles different data types (numerics, factors, etc), and returns a skimr object that can be piped or displayed nicely for the human reader. 

## Installation


``` r
# install.packages("devtools")
devtools::install_github("ropenscilabs/skimr")
```

To install the version with the most recent changes that have not yet been incorporated in 
the master branch (and may not be):

``` r
devtools::install_github("ropenscilabs/skimr", ref = "develop")
```

## Skim statistics in the console

- added missing, complete, n, sd
- reports numeric/int/double separately from factor/chr
- handles dates, logicals
- supports spark-bar and spark-line based on 
[Hadley Wickham's pillar package](https://github.com/hadley/pillar).

**Nicely separates variables by class:**  

```{r}
skim(chickwts)
```

**Presentation is in a compact horizontal format:**  
```{r}
skim(iris)
```

Individual columns of a data frame can be selected using tidyverse style selectors.

```{r}
skim(iris, Sepal.Length, Petal.Length)
```

## Handles grouped data

Skim() can handle data that has been grouped using `dplyr::group_by`.

```{r}
iris %>% dplyr::group_by(Species) %>% skim()
```

## Options for kable and pander

Enhanced print options are available by piping to kable() or pander(). 

## skim_df object (long format)

By default `skim` prints beautifully in the console, but it also produces a long, tidy-format skim_df object that can be computed on. 

```{r}
a <-  skim(chickwts)
dim(a)
```

```{r}
print.data.frame(skim(chickwts))
```

## Compute on the full skim_df object
```{r}
skim(mtcars) %>% dplyr::filter(stat=="hist")
```

## Works with strings, lists and other column classes.

```{r}
skim(dplyr::starwars)
```

Users can add new classes.

## Specify your own statistics

```{r}
funs <- list(iqr = IQR,
    quantile = purrr::partial(quantile, probs = .99))
  skim_with(numeric = funs, append = FALSE)
  skim(iris, Sepal.Length)
# Restore defaults
  skim_with_defaults()
```

## Limitations of current version

We are aware that there are issues with rendering the inline histograms and line charts in various contexts, some of which are 
described below. 

### Windows support for spark histograms

Windows cannot print the spark-histogram characters when printing a data-frame. For example, 
`"▂▅▇"` is printed as `"<U+2582><U+2585><U+2587>"`. This longstanding problem [originates in 
the low-level code](http://r.789695.n4.nabble.com/Unicode-display-problem-with-data-frames-under-Windows-td4707639.html) 
for printing dataframes.

This means that while `skimr` can render the histograms to the console and in
kable, it cannot in other circumstances. This includes:

* rendering a `skimr` data frame within pandaer
* converting a `skimr` data frame to a vanilla R data frame, but tibbles render
  correctly

In previous versions of One workaround for showing these characters in Windows is to set the CTYPE part of your locale to Chinese/Japanese/Korean with `Sys.setlocale("LC_CTYPE", "Chinese")`. These values do show up by default when printing a data-frame created by `skim()` as a list (`as.list()`) or as a matrix (`as.matrix()`).

### Printing spark histograms and line graphs in knitted documents

Spark-bar and spark-line work in the console but may not work when you knit them to a specific document format.
The same session that produces a correctly rendered HTML document may produce an incorrectly rendered PDF, 
for example. This issue can generally be addressed by changing fonts to one with good building block (for
histograms) and braille support (for line graphs).  For example, the open font "DejaVu Sans" from 
the `extra font` package supports these.  You may also want to try wrapping your results in `knitr::kable()`.
Please see the vignette on using fonts for details on this.

Displays in documents of different types will vary. For example, one user found that the font 
"Yu Gothic UI Semilight"  produced consistent results for Microsoft Word and Libre Office Write.

## Contributing

We welcome issue reports and pull requests including potentially adding support for different variable classes. Please see the contributing.md document.
